\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-07-27 popl-2019-journal-v/proving-full-abstraction.tex
Readd a comment that got lost in the merge...
[POPLjv] Correct later cofe and preordered cofe definition
[POPLjv] Update terminology for regions to spatial, shadow, and shared

For the sake of readability, we introduce the following notation
\begin{gather*}
  \pwheap = \pi_1(W)\\
  \pwpriv = \pi_2(W)\\
  \pwfree = \pi_3(W)
\end{gather*}
\subsubsection{Linearity}
% Linearity
The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
In order to model this uniqueness, we need to keep track of which parts of memory are uniquely referenced and make sure that only one linear capability references the unique parts.
%% Bookkeeping done by worlds
We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
If a memory segment is governed by a shared region, then normal capabilities may reference it.
On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.
%%% spatial/spatial_owned
\todo[inline]{Dominique: Maybe it would be clearer to rename spatial\_owned/spatial to spatial/spatial\_shadow or something similar. Spatial\_owned is really the default meaning of spatial, and what we've called spatial is a peculiar thing that we use for a very specific purpose.
  This would make the discussion easier to follow, I think.}

We cannot let multiple linear capabilities reference the same memory, so we add ownership to spatial regions.
Only the spatial owned regions can be referenced by a capability.
Even though only spatial owned regions can be referenced by linear capabilities, we still have unclaimed spatial regions that specify what part of memory looks like.
When there is an unclaimed spatial region, then there may be a compatible world  that claims the ownership of that spatial region which means that the memory may still be referenced (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatial$ and $\spatialo$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
  \end{array} \right.
\]
\todo[inline]{Dominique: Perhaps rename pure to shared?  Or introduce the term pure earlier on?}
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

@@

We refer to this as the pure part of the world and define the function $\purePart{}$ which removes all ownership.
\begin{definition}[The pure part of a world]

@@

  \todo[inline]{Maybe explain here that Kripke future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
    However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on the capability anymore.
    By linking revocation of regions immediately to linearity, it will require less explanation later that only spatial regions can be revoked.
  }
In order to relate \trgcm{} and \srccm{}, we are mainly interested in changes to the stack as it repurposes stack frames.
Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions for call frames that are popped from the call stack, to make place for a new frame.
We allow introduction of new regions simply by extension.
That is, a future world can have new regions that was not present in a past world.
Things are a bit more complicated when it comes to revocation.
\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).
}
\dominique{To be honest, I don't actually know what exactly would go wrong if we simply removed revoked regions.
  Perhaps we should think this through?
}
For technical reasons, we need the worlds to be extensional\todo{Dominique: meaning?}, so we cannot simply forget about regions that we want to remove.
We keep removed regions in the world and tag them as removed with a $\revoked$ tag.
This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.

@@

  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
  \end{align*}
\end{definition}
% TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
Functions may not only retain equivalences but make it more precise.
Such a function is said to be \emph{contractive}.
\begin{definition}
  \label{def:nonexpansive-contractive-ofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
  The function $f$ is
  \begin{description}
  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal_Y f(x')
\]
  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
  \qedhere
\]
  \end{description}
\end{definition}
\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
There are a number of standard c.o.f.e.\ constructions for countably infinite sets, products, sums, and functions.
A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
The other constructions are used to create c.o.f.e.'s from other c.o.f.e.'s.
For instance the product construction is defined as follows:
\begin{definition}[Product c.o.f.e.]
\label{def:product-cofe}
Given two c.o.f.e.'s $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal \right)_{n=0}^{\infty} \right)$ where the equivalence family is defined as for $(x,y),(x',y') \in X \times Y$
\[
  (x,y) \nequal (x',y') \text{ iff } x \nequal x' \wedge y \nequal y'
\]
\end{definition}
The remaining standard constructions are defined similarly.
Another important construction is the \emph{later} ($\blater$) c.o.f.e.\
The later c.o.f.e.\ construction is based on one c.o.f.e.\ which it decreases the index of.
\begin{lemma}[$\blater$ c.o.f.e.]
  \label{def:later-cofe}
  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
  \]
  where
  \[
    x \nequal_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{}\\
        x \nequal[n+1]_X x'
      \end{array}
    \right.
  \]
\end{lemma}
The later c.o.f.e.\ is used to make sure that the index decreases before the self-reference which gives something to recurse on in order to find a solution for the domain equation.
If we only use the standard constructions to construct a self-referential equation and make sure to guard every negative occurrence of the self reference with the later c.o.f.e.\ construction, then the equation has a solution.
\todo[inline]{Insert the solution theorem here?}

Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
The worlds we sketched have additional structure in terms of a future world relation.
The future world relation is captured by adding a preorder to c.o.f.e.'s:
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
We will need the standard construction for later, product, and sum later, so we define them here.
The reader may skip these definitions for now and refer back to them later.
\begin{definition}[$\blater$ preordered c.o.f.e.]
  \label{def:later-p-cofe}
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
  \]
  where
  \[
    x \nequal_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{} \\
        x \nequal
        [n+1]_X x'
      \end{array}
    \right.
  \]
\end{definition}
We will also need the canonical preordered c.o.f.e.\ for sum.
\begin{definition}[Product preordered c.o.f.e.]
  \label{def:product-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
\end{definition}
\begin{definition}[Sum preordered c.o.f.e.]
  \label{def:sum-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X + Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
  \[
    \text{for } z,z' \in X + Y,\quad  z' \future z \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z' \future_x z \vee{} \\
        z,z' \in Y \wedge z' \future_Y z
      \end{array}
    \right.
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
    \text{for } z,z' \in X + Y,\quad  z \nequal z' \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z \nequal_x z' \vee{} \\
        z,z' \in Y \wedge z \nequal_Y z'
      \end{array}
    \right.
  \]
\end{definition}

@@

Next, we want to define the regions of the world.
As mentioned, regions have the added structure of a future region relation, so we need to define them as a preordered c.o.f.e.
The future region relation was presented in Section~\ref{subsubsec:ft-and-revocation}, so we just need to use it for our definitions.
The future world relation for pure region is trivial as it is just equality
The pure regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
\begin{definition}[Pure region preordered c.o.f.e.\ ]
  \label{def:pure-reg-p-cofe}
  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of pure regions as
  \[
  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
  \]
  where the preorder $\future$ is equality, and the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
  \[
  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
  \]
\end{definition}
Definition~\ref{def:pure-reg-p-cofe} gives us the pure region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.

@@

Finally, we define the sub-world for the heap.
To this end, we combine the pure region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
The result is a preordered c.o.f.e.\
\[
  \Regions + \Regionh
\]
which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
\[
  \Worldh = \nats \parfun (\Regions + \Regionh)
\]

@@

The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
We later define the $\lrename$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
An executable capability can have its range of authority shrunk or its current address changed which changes what instructions are executed and thus potentially whether the code respects memory invariants.
For this reason, the condition requires that any executable capability with a derived range of authority and a current address in that range is safe to use for execution.
The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$.
We do not know when the executable capability will be used, so it should be safe even in the future when the memory has changed.
The $\purePart{}$ function strips the world of all spatial ownership.
This means that the capability cannot depend on linear capabilities and thus the contents of the stack.
When we define the logical relation, we even require the executable capability to not be linear.
Linear executable capabilities would likely not be useful because they cannot be moved from the $\pcreg$-register without crashing the execution.
This may sound like an ideal primitive for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets using the program counter capability which is not possible when the program counter is linear.
\begin{align*}
  \execCond{W} &=
  \left\{ \npair{A} \middle|
    \begin{array}{l}
      \forall n' < n, W' \future \purePart{W}\ldotp \forall b',e'\ldotp \forall \aaddr \in [b',e'] \subseteq A \ldotp\\
      \quad \npair[n']{\left( \arraycolsep=0pt\array{l}((\rx,\normal),\baddr',\eaddr',\aaddr),\\((\rx,\normal),\baddr',\eaddr',\aaddr)\endarray\right)} \in \lre(W')
    \end{array}
    \right\}
\end{align*}
The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memory segments.
However, when we have a read capability with execute permission, we know that the capability must point to a piece of code memory.
For this reason, we define the $\xReadCond{}$ to capture the additional assumptions that we can make when a capability is executable.
\[
  \xReadCond{W} = \left\{ \npair{A} \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\normal, \pwheap}, \code \ldotp \\
      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
      \qquad \dom(\code) \supseteq A 
    \end{array}
  \right\}
\]
The $\xReadCond{}$ requires that the memory segment an executable capability has authority over is governed by a code region.
Note that we do not define $\execCond{}$ and $\xReadCond{}$ for the stack because the stack is not executable.

>>>>

For the sake of readability, we introduce the following notation
\begin{gather*}
  \pwheap = \pi_1(W)\\
  \pwpriv = \pi_2(W)\\
  \pwfree = \pi_3(W)
\end{gather*}
\subsubsection{Linearity}
% Linearity
The linear capabilities of \srccm{} and \trgcm{} guarantee that they have the sole authority over the memory they reference\footnote{Under the assumption that the system was initialised with unique linear capabilties.}.
In order to model this uniqueness, we need to keep track of which parts of memory are uniquely referenced and make sure that only one linear capability references the unique parts.
%% Bookkeeping done by worlds
We use the world to keep track of what parts of memory must be uniquely referenced by having two kinds of regions: shared and spatial.
If a memory segment is governed by a shared region, then normal capabilities may reference it.
On the other hand, if a memory segment is governed by a spatial region, then only a linear capability may reference it.

%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory.
The spatial region represent ownership a part of memory which ensures that a linear capability is not aliased.
Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
To this end, we have shadow regions.
A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
This does not mean that the memory is necessarily not referenced.
A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

@@

We refer to this as the shared part of the world and define the function $\purePart{}$ which turns all spatial regions into shadow copies.
\begin{definition}[The shared part of a world]

@@

  \todo[inline]{Maybe explain here that Kripke future world relations usually only allow for adding fresh assumptions on fresh memory or protocol steps that the system was designed to support from the start.
    However in our setting, we sometimes allow dropping assumptions, namely when linearity tells us that no value in the system depends on the capability anymore.
    By linking revocation of regions immediately to linearity, it will require less explanation later that only shadow regions can be revoked.
  }
In order to relate \trgcm{} and \srccm{}, we are mainly interested in changes to the stack as it repurposes stack frames.
Specifically, we need to introduce regions to govern new encapsulated stack frames, and we need to revoke regions for call frames that are popped from the call stack, to make place for a new frame.
We allow introduction of new regions simply by extension.
That is, a future world can have new regions that was not present in a past world.
Things are a bit more complicated when it comes to revocation.
\lau{Can we give a more precise explanation of why regions cannot be removed from the world (still in terms of intuition).
}
\dominique{To be honest, I don't actually know what exactly would go wrong if we simply removed revoked regions.
  Perhaps we should think this through?
}
For technical reasons, we need the worlds to be extensional\todo{Dominique: meaning?}, so we cannot simply forget about regions that we want to remove.
We keep removed regions in the world and tag them as removed with a $\revoked$ tag.
This means that the regions should be able to get the $\revoked$ tag which we express in a future region relation.

@@

  An element $x \in X$ is the \emph{limit} of the sequence $\seq{x}$ if
  \begin{align*}
    \forall k \in \nats, \exists j \in \nats, \forall n \geq j, x \nequal[k] x_n.
  \end{align*}
\end{definition}
% TODO write something here about why the different things (the meassure of the ofe and the guaranteed limits of the cofe) gives us in terms of finding a solution.
In order to move from one c.o.f.e.\ to another, we define functions, just like we would when working with sets.
Unlike functions between sets, functions between c.o.f.e.'s must preserve the c.o.f.e.\ structure which informally means that the measure of precision must be retained.
This specifically means that functions between c.o.f.e.'s must be \emph{non-expansive} which basically means that they preserve $n$-equalities.
Functions may not only retain equivalences but make it more precise.
Such a function is said to be \emph{contractive}.
\begin{definition}
  \label{def:nonexpansive-contractive-ofe}
  Let $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty} \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty} \right)$ be two ordered families of equivalences and $f$ a function from the set $X$ to the set $Y$.
  The function $f$ is
  \begin{description}
  \item[non-expansive] when for all $x, x' \in X$, and all $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal_Y f(x')
\]
  \item[contractive] when for any $x, x' \in X$, and any $n \in \nats$,
\[
  x \nequal_X x' \implies f(x) \nequal[n+1]_Y f(x') 
  \qedhere
\]
  \end{description}
\end{definition}
\noindent The step-indexing provided by c.o.f.e.'s give enough structure to solve the recursive domain equation.
There are a number of standard c.o.f.e.\ constructions for countably infinite sets, products, sums, and functions.
A set can be lifted to c.o.f.e.\ by having the normal equality for all positive indices and the total relation for $\nequal[0]$.
The other constructions are used to create c.o.f.e.'s from other c.o.f.e.'s.
For instance the product construction is defined as follows:
\begin{definition}[Product c.o.f.e.]
\label{def:product-cofe}
Given two c.o.f.e.'s $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ and $\left(Y, \left(\nequal_Y\right)_{n=0}^{\infty} \right)$ define the product c.o.f.e.\ as $\left(X \times Y, \left(\nequal \right)_{n=0}^{\infty} \right)$ where the equivalence family is defined as for $(x,y),(x',y') \in X \times Y$
\[
  (x,y) \nequal (x',y') \text{ iff } x \nequal x' \wedge y \nequal y'
\]
\end{definition}
The remaining standard constructions are defined similarly.
Another important construction is the \emph{later} ($\blater$) c.o.f.e.\
The later c.o.f.e.\ construction is based on one c.o.f.e.\ which it decreases the index of.
\begin{lemma}[$\blater$ c.o.f.e.]
  \label{def:later-cofe}
  Given a c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right)$ define the later c.o.f.e.\ as
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty} \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty} \right)
  \]
  where
  \[
    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{}\\
        x \nequal_X x'
      \end{array}
    \right.
  \]
\end{lemma}
The later c.o.f.e.\ is used to make sure that the index decreases before the self-reference which gives something to recurse on in order to find a solution for the domain equation.
If we only use the standard constructions to construct a self-referential equation and make sure to guard every negative occurrence of the self reference with the later c.o.f.e.\ construction, then the equation has a solution.
\todo[inline]{Insert the solution theorem here?}

Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
The worlds we sketched have additional structure in terms of a future world relation.
The future world relation is captured by adding a preorder to c.o.f.e.'s:
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
We will need the standard construction for later, product, and sum later, so we define them here.
The reader may skip these definitions for now and refer back to them later.
\begin{definition}[$\blater$ preordered c.o.f.e.]
  \label{def:later-p-cofe}
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
  \]
  where
  \[
    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{} \\
        x \nequal
        [n]_X x'
      \end{array}
    \right.
  \]
\end{definition}
We will also need the canonical preordered c.o.f.e.\ for sum.
\begin{definition}[Product preordered c.o.f.e.]
  \label{def:product-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
\end{definition}
\begin{definition}[Sum preordered c.o.f.e.]
  \label{def:sum-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X + Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
  \[
    \text{for } z,z' \in X + Y,\quad  z' \future z \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z' \future_x z \vee{} \\
        z,z' \in Y \wedge z' \future_Y z
      \end{array}
    \right.
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
    \text{for } z,z' \in X + Y,\quad  z \nequal z' \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z \nequal_x z' \vee{} \\
        z,z' \in Y \wedge z \nequal_Y z'
      \end{array}
    \right.
  \]
\end{definition}

@@

Next, we want to define the regions of the world.
As mentioned, regions have the added structure of a future region relation, so we need to define them as a preordered c.o.f.e.
The future region relation was presented in Section~\ref{subsubsec:ft-and-revocation}, so we just need to use it for our definitions.
The future world relation for the shared region is trivial as it is just equality
The shared regions we sketched are the easiest to define in terms of a c.o.f.e.\ as the future relation is just equality.
\begin{definition}[Shared region preordered c.o.f.e.\ ]
  \label{def:pure-reg-p-cofe}
  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{SealInv},\left(\nequal[n]_{\var{SI}}\right)_{n=0}^{\infty}\right)$, define the preordered c.o.f.e.\ of shared regions as
  \[
  \left(\{\pure\}\times \var{MemInv} \times \var{SealInv},\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
  \]
  where the preorder $\future$ is equality, and the family of equivalences distributes to the underlying c.o.f.e.'s, i.e. given $(\pure,H,H_\sigma), (\pure,H',H_\sigma') \in \{\pure\}\times \var{MemInv} \times \var{SealInv}$,
  \[
  (\pure,H,H_\sigma) \nequal (\pure,H',H_\sigma') \text{ iff } H \nequal_\var{MI} H' \wedge H_\sigma \nequal_{SI} H_\sigma'
  \]
\end{definition}
Definition~\ref{def:pure-reg-p-cofe} gives us the shared region for preorder c.o.f.e.\ when we use it with the memory invariant c.o.f.e.\ as $\var{MemInv}$ and the seal invariant c.o.f.e.\ as $\var{SealInv}$.

@@

Finally, we define the sub-world for the heap.
To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
The result is a preordered c.o.f.e.\
\[
  \Regions + \Regionh
\]
which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
\[
  \Worldh = \nats \parfun (\Regions + \Regionh)
\]

@@

The $\execCond{}$ intuitively says that an execute capability is safe when any capability that can be derived from it is safe as a program counter now and in the future.
We later define the $\lrename$-relation which captures what it means for a word to be safe as a program counter, but for now it suffices to think of it as a program counter that causes an execution that does not break memory invariants.
An executable capability can have its range of authority shrunk or its current address changed which changes what instructions are executed and thus potentially whether the code respects memory invariants.
For this reason, the condition requires that any executable capability with a derived range of authority and a current address in that range is safe to use for execution.
The $\execCond{}$ is quantified over all future worlds of the $\purePart{}$ of $W$.
We do not know when the executable capability will be used, so it should be safe even in the future when the memory has changed.
The $\purePart{}$ function turns the spatial regions of a world into shadow copies.
This means that the capability cannot depend on linear capabilities and thus the contents of the stack.
When we define the logical relation, we even require the executable capability to not be linear.
Linear executable capabilities would likely not be useful because they cannot be moved from the $\pcreg$-register without crashing the execution.
This may sound like an ideal primitive for constructing something that can be executed once, however, most programs rely on loading other capabilities or seal sets using the program counter capability which is not possible when the program counter is linear.
\begin{align*}
  \execCond{W} &=
  \left\{ \npair{A} \middle|
    \begin{array}{l}
      \forall n' < n, W' \future \purePart{W}\ldotp \forall b',e'\ldotp \forall \aaddr \in [b',e'] \subseteq A \ldotp\\
      \quad \npair[n']{\left( \arraycolsep=0pt\array{l}((\rx,\normal),\baddr',\eaddr',\aaddr),\\((\rx,\normal),\baddr',\eaddr',\aaddr)\endarray\right)} \in \lre(W')
    \end{array}
    \right\}
\end{align*}
The $\readCond{}$ condition by itself allows many different regions and thus potentially many different memory segments.
However, when we have a read capability with execute permission, we know that the capability must point to a piece of code memory.
For this reason, we define the $\xReadCond{}$ to capture the additional assumptions that we can make when a capability is executable.
\[
  \xReadCond{W} = \left\{ \npair{A} \middle| 
    \begin{array}{l}
      \exists r \in \addressable{\normal, \pwheap}, \code \ldotp \\
      \qquad \pwheap(r) \nequal \codereg{\_,\_,\code,\gc}\wedge\\
      \qquad \dom(\code) \supseteq A 
    \end{array}
  \right\}
\]
The $\xReadCond{}$ requires that the memory segment an executable capability has authority over is governed by a code region.
Note that we do not define $\execCond{}$ and $\xReadCond{}$ for the stack because the stack is not executable.
